Leadership across cultures and contexts: an exploratory analysis

### 2020-04-15 leadership_across_cultures_contexts.Rmd
Minor edits across manuscript.

We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations [i.e., @henrich_weirdest_2010]. We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations. Through novel exploratory analyses encompassing the full range of cultural diversity and social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.

@@

Leaders emerge in all human societies. That is, within every population there are particular contexts in which one individual maintains a disproportionate level of influence in the group and functions to facilitate group goals and provides services to the group. Adopting the broad definition of leadership given above, parents can also be defined as leaders of nuclear families. Strong community leadership, however, is extremely common cross-culturally including among hunter-gatherers and relatively small-scale societies [@garfield_review_2019; @von_rueden_leadership_2015].

Despite the universality of the phenomenon of leadership, the nature of leadership across populations and across social contexts is highly diverse and the emergence of leaders and the functions they serve are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007]. Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015]. In the absence of institutionalized organizational structure leaders often take on essential roles in group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who do not appropriately contribute to collective actions, comply with group policies, or seriously violate social norms. @glowacki_leadership_2015 draw on empirical, ethnographic, and theoretical models of leader emergence and suggest in small-scale societies institutionalized leadership roles are likely to initially emerge to resolve conflicts between families. @roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military fromidability to rival groups and facilitating within-group conflict resolution and cooperation. In addition to military prowess, 'big-men' are characterized by, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p. 104].

Given specific demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993]. The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019]. Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008]. Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012]. Across relatively small-scale societies, at the community level, representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of leaders. Within communities, however, leaders may emerge whenever individuals form groups. The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions leaders serve.

@@

In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990]. Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005]. These positions typically are associated with institutionalized responsibilities and specific functions associated with their roles. The structure of such organizations are generally organized hierarchically with nested bodies and a formal internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010]. They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.

@@

More recently, many managerial scholars have adopted integrative and comparative approaches to better understand cultural diversity in leadership systems. @behrendt_integrative_2017 draw on empirical results from the managerial literature and suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources. This framework is intended to provide a more fine-grained and externally valid view of leadership functions, compared to traditional models of transactional leadership, which tend to oversimplify dimensions of leader-follower dynamics. More integrated and holistic models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning. Culturally Endorsed Implicit Leadership Theory (CLT), suggests cultural variation in social norms has the potential to impact follower preferences for leadership styles [@moore_literature_2018]. The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leader effectiveness and cultural variation across large organizations. The GLOBE project has produced multiple cross-national data sets on leadership and hundreds of publications. For example, *The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* draws on data from over 1,000 CEOs and over 5,000 senior executives from corporations in varied of industries across 24 countries [@house2013strategic]. The goals of the GLOBE CEO Study were to address (1) how national culture influences expectations for the behaviors of leaders, (2) which behaviors of CEOs are typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations. Through path analysis the study concluded that cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations. That is, leaders moderate their behavior through culturally learned values and leadership-specific values.

@@

<!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
In contrasting the functions leaders serve in smaller-scale, non-industrial, and more traditional societies (i.e., non-WEIRD), from larger-scale, (post)industrial contexts (i.e., WEIRD), community leaders in non-WEIRD settings are more likely to have domain-specific functions based on role-specialization [see @smith_leadership_2016]. The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies. In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple domains. In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations. Community leadership functions in increasingly WEIRD contexts will likely be increasingly focused on economics and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage inter-group social dynamics. 
<!-- Could add ethnographic examples for SSS and LSS  -->

@@

The type of group to which a leader text referred was coded into seven researcher-coded categories: residential groups, kin groups, economic groups, political groups at the community level, political group beyond the community level, military groups, and religious groups. Figure \@ref(fig:grouptype) indicates the proportion of text records that were categorized into each of these groups, broken out by mode of subsistence. We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category. We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.

@@

We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)). Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)`. There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors). Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author. Total pages of ethnography on each culture in the PSF ranged from `r min(dfwc[['N Pages eHRAF']])` to `r as.integer(max(dfwc[['N Pages eHRAF']]))`, with a median of `r median(dfwc[['N Pages eHRAF']])`.

We fit `r length(all_study_vars)` logistic mixed effects models with each binary study variable as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures. We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995]. After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models. More recent publications were more likely to report evidence for Oratory skill, Group representative, Group determination/cohesiveness, and Knowledgeable/intelligent, and less likely to report evidence for leaders imposing costs (Punishment, Killer) or receiving benefits (Social services, Food, and Reduced risk of harm). See Figure \@ref(fig:pubdate). See supplementary information for details.

@@

Our qualities and functions data comprised sparse $`r nrow(leader_text)` \times `r length(quality_vars)`$ and $`r nrow(leader_text)` \times `r length(function_vars)`$ matrices of binary data, respectively.

@@

This exploratory study extracted the qualities, functions, benefits, and costs of leadership from the ethnographic record. These variables were *not* determined prior to the study but were instead identified and coded during several passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section). Our results therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers across a diverse and culturally representative sample of the ethnographic record. These results reveal universal dimensions of human leadership while also suggesting important variation by group type, subsistence type, and sex of the leader.

@@

We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful categories. Variables were clustered together if they were positively correlated. Leader qualities formed two top-level clusters, which we identify as a *Material, social, and reproductive success* cluster, and a *Social orientation* cluster. See Figure \@ref(fig:cluster)A. Several of traits often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:

@@

The evolution of leadership and followership is somewhat perplexing. Leadership often involves the provisioning of a public good at an individual cost, raising the question of how the individual benefits of leadership outweigh the costs. Followership, on the other hand, involves relinquishing individual decision-making to a leader who might or might not act in the follower's interest. The benefits of followership (and by extension, group living) must also outweigh the costs for leadership to evolve and be maintained. We therefore coded all benefits and costs of both leadership and followership to better understand which benefits and costs are most commonly associated with leading or following. Leaders were widely reported to receive material benefits and social status, but loss of material resources and social status were also widely reported. The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources. Our results cannot determine if the benefits to leaders and followers outweighed their costs, but they do shed light on the social and biological pressures associated with leadership and social hierarchy. For both leaders and followers, individuals stood to lose what they were able to gain. The major difference between leaders and followers lay in leaders' opportunities for increased social status and mating opportunities, relative to followers. See Figure \@ref(fig:varsupportcb).

@@

```{r eHRAFpages, fig.height=6, fig.cap="Scatterplot of the number of leadership text records versus the total number of pages of ethnography in the eHRAF. Each dot is one culture. Line fit by loess regression."}
plot_pages_tr
```

>>>>

We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations [i.e., @henrich_weirdest_2010]. We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations. Through novel exploratory analyses encompassing the full range of cultural diversity and diverse social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.

@@

Leaders emerge in all human societies. That is, within every population there are particular contexts in which one individual maintains a disproportionate level of influence in the group, functions to facilitate group goals, and provides services to the group. Adopting the broad definition of leadership given above, parents can also be defined as leaders of nuclear families. Strong community leadership, however, is extremely common cross-culturally including among hunter-gatherers and relatively small-scale societies [@garfield_review_2019; @von_rueden_leadership_2015].

Despite the universality of the phenomenon of leadership, the nature of leadership across populations and across social contexts is highly diverse and the emergence of leaders and the functions they serve are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007]. Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015]. In the absence of institutionalized organizational structure leaders often take on essential roles in group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who do not appropriately contribute to collective actions, comply with group policies, or seriously violate social norms. @glowacki_leadership_2015 draw on empirical, ethnographic, and theoretical models of leader emergence and suggest in small-scale societies institutionalized leadership roles are likely to initially emerge to resolve conflicts between families. @roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military fromidability to rival groups and facilitating within-group conflict resolution and cooperation. For "big-men" in these societies, in addition to military prowess, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p. 104].

Given demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993]. The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019]. Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008]. Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012]. Across relatively small-scale societies representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of community leaders. Within communities, however, leaders may emerge whenever individuals congregate. The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions carried out by a leader.

@@

In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990]. Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders, to name a few [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005]. These positions typically are associated with institutionalized responsibilities and specific functions. The structure of such organizations are generally hierarchical with nested bodies and an internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010]. They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.

@@

More recently, many managerial scholars have adopted integrative and comparative approaches to better understand cultural diversity in leadership systems. @behrendt_integrative_2017 draw on empirical results from the managerial literature and suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources. This framework is intended to provide a more fine-grained and externally valid view of leadership functions, compared to traditional models of transactional leadership, which tend to oversimplify dimensions of leader-follower dynamics. More integrated and holistic models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning. Culturally Endorsed Implicit Leadership Theory (CLT), suggests cultural variation in social norms has the potential to impact follower preferences for leadership styles [@moore_literature_2018]. The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leader effectiveness and cultural variation across large organizations. The GLOBE project has produced multiple cross-national data sets on leadership and hundreds of publications. For example, *The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* draws on data from over 1,000 CEOs and over 5,000 senior executives from corporations in varied industries across 24 countries [@house2013strategic]. The goals of the GLOBE CEO Study were to address (1) how national culture influences expectations for the behaviors of leaders, (2) which behaviors of CEOs are typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations. Through path analysis the study concluded that cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations. That is, leaders moderate their behavior through culturally learned values and leadership-specific values.

@@

<!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
In contrasting the functions leaders serve in smaller-scale, non-industrial, and more traditional societies (i.e., non-WEIRD), from larger-scale, (post)industrial contexts (i.e., WEIRD), community leaders in non-WEIRD settings are more likely to have domain-specific functions based on role-specialization [see @smith_leadership_2016]. The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies. In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple domains. In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations. Community leadership functions in increasingly WEIRD contexts will likely be increasingly focused on economics and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage intra-group social dynamics.
<!-- Could add ethnographic examples for SSS and LSS  -->

@@

The type of group to which a leader text referred was coded into seven researcher-coded categories: residential groups, kin groups, economic groups, political groups at the community level, political group beyond the community level, military groups, and religious groups (see Methods section). Figure \@ref(fig:grouptype) indicates the proportion of text records that were categorized into each of these groups, broken out by mode of subsistence. We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category. We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.

@@

We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)). Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure \@ref(fig:doc-years)). There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors). Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author. Total pages of ethnography on each culture in the PSF ranged from `r min(dfwc[['N Pages eHRAF']])` to `r as.integer(max(dfwc[['N Pages eHRAF']]))`, with a median of `r median(dfwc[['N Pages eHRAF']])`.

We fit `r length(all_study_vars)` logistic mixed effects models with each binary study variable as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures. We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995]. After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models. More recent publications were more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*). See Figure \@ref(fig:pubdate). See supplementary information for details.

@@

Our qualities and functions data comprised sparse $`r nrow(leader_text2)` \times `r length(quality_vars)`$ and $`r nrow(leader_text2)` \times `r length(function_vars)`$ matrices of binary data, respectively.

@@

This exploratory study extracted the qualities, functions, benefits, and costs of leadership from the ethnographic record. The variables on leader qualities and functions were *not* determined prior to the study but were instead identified and coded during several passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section). Our results therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers across a diverse and culturally representative sample of the ethnographic record. These results reveal universal dimensions of human leadership while also suggesting important variation by group type, subsistence type, and sex of the leader.

@@

We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful categories. Variables were clustered together if they were positively correlated. Leader qualities formed two top-level clusters, which we identify as a *Material, social, and reproductive success* cluster, and a *Social orientation* cluster. See Figure \@ref(fig:cluster)A. Several of these traits often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:

@@

The evolution of leadership and followership is somewhat perplexing. Leadership often involves the provisioning of a public good at an individual cost, raising the question of how the individual benefits of leadership outweigh the costs. Followership, on the other hand, involves relinquishing individual decision-making to a leader who might or might not act in the follower's interest. The benefits of followership (and by extension, group living) must also outweigh the costs for leadership to evolve and be maintained. We therefore coded all benefits and costs of both leadership and followership to better understand which benefits and costs are most commonly associated with leading or following. Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported. The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources. Followers were not reported to recieve mating benefits, relative to leaders or other types benefits. Our results cannot determine if the benefits to leaders and followers outweighed their costs, but they do shed light on the social and biological pressures associated with leadership and social hierarchy. For both leaders and followers, individuals stood to lose what they were able to gain. The major difference between leaders and followers lay in leaders' opportunities for increased social status and mating opportunities, relative to followers. See Figure \@ref(fig:varsupportcb).

@@

## Ethnographic sample

```{r}
doc_1950_perc <- signif(length(documents$d_publication_date[documents$d_publication_date>1949])/length(documents$d_publication_date),2)*100
```

The majority of ethnographic documents sampled here (`r doc_1950_perc`%) were published after 1950 and the mean publication year is `r round(mean(documents$d_publication_date),)` (SD = `r round(sd(documents$d_publication_date),2)`). Figure \@ref(fig:doc-years) plots the publication years of all documents which contributed text records to the ethnographic sample.

```{r doc-years, fig.height=6, fig.cap="Historgram of eHRAF document publiction year. Vertical line shows the mean publication year."}
doc_year_hist
```

The eHRAF provides information on the time frame of field work associated with each document. For some documents the span of associated fieldwork is given for the exact months and years the ethnographer was in the field. For most documents one or more spans of years are given. Figure \@ref(fig:field-years) plots the earliest year and the latest year for which field work occurred for each document. The time spans plotted do not depict the full span of fieldwork necessarily, but simply illustrate the time span during which fieldwork occurred. We omitted data for three documents (two from North America and one from Insular Pacific) with field work date ranges prior to 1870. For all documents, the average earliest year in which field work occurred was `r signif(mean(documents$d_field_date_start, na.rm=T),4)` and the average latest years in which field work occurred was `r signif(mean(documents$d_field_date_end, na.rm=T),4)`.

```{r field-years, fig.height=8, fig.width=9, fig.cap="Dumbbell plot of eHRAF document field work years. Vertical lines shows the  mean earliest and latest years of field work."}
doc_fielddates_plot
```

```{r eHRAFpages, fig.height=8, fig.cap="Scatterplot of the number of leadership text records versus the total number of pages of ethnography in the eHRAF. Each dot is one culture. Line fit by loess regression."}
plot_pages_tr
```
